Article Title

### 2023-10-25 response_to_review.Rmd

---
title: "Response to review"
---

**Overview**

Thank you to the editor and reviews for your helpful feedback on our paper. We have made the changes requested, and explain them below, in *italicised* text.

**Reviewer #1**

This paper proposes an interactive visualization tool and its implementation as an R package to intuitively understand LVA in non-linear models.

General Comments

You have not clearly stated the usefulness of the radial tour; please explain the advantage of using the SHAP values of PI and CI as the contribution of 1D projection compared to drawing a static density plot or scatterplot. Is it possible to correctly evaluate the sensitivity of variables by 1d projection for nonlinear models? Also, in the Case Studies in Section 5, please provide an explanation in relation to the Radial tour operation.

*TODO (DC):*

The results are very different between the examples provided in the shiny application in the R cheem package and the examples presented in the paper. Please make sure to use random number seeds to achieve similar results. In particular, in the Case Studies of classification, there is no misclassification in the examples in the cheem package, and the misclassification is not indicated by the red circle as shown in the paper.

*TODO (NS):*

The panel that is treeshap in the global view in the paper is attribution in the cheem package. Please unify either one.

*TODO (NS):*

In the global view in the paper, the model panel has jitter processing, but the cheem package does not.

*TODO (NS):*

Discussion is just a Conclusion. For example, please supplement any discussion on the following.
- Abstract and Introduction are described in the context of XAI, but the tools proposed in this paper are not likely to be able to handle large data sets.
- Concerning the convenience of the package: the application by the cheem package outputs little information (e.g., axis information), which makes it difficult to use intuitively.

*TODO (DC):*

Individual Comments

(p.7 l.18) Figures 3b, 4a -> 3c, 4b?

*TODO (DC):*

(p.8 l.7) Is the SHAP value of the CI taken into account in the radial tour?

*TODO (DC):*

(p.8 l.8) "The PI is ... global view" : The shiny app in the cheem package does not support interactive selection by clicking in the global view. (cheem version 0.3.0 on Windows 11 / Google Chrome)

*TODO (PB):*

(p.10 Fig4) Residual plot is not displayed in cheem version 0.3.0 on Windows 11 / Google Chrome.

*TODO (PB):*

(p.11 l.5) Figure 3a & b -> 4a & b?

*TODO (PB):*

(p.11 l.7) Figure 3d -> 4d?

*TODO (PB):*

(p.21 l.10) https://nspyrison.github.io/cheem/ is not available.

*TODO (NS): And Nick, if you need help we will need to do a zoom call. I think that I mentioned how to fix this some time ago, but not sure you understood.*

(references) There are some characters like "???".

*TODO (PB):*

**Reviewer #2:**

I was able to run your software using publicly available packages.
Unfortunately, I could not play videos shown at the following URLs.

*We cannot address this unless the error messages are provided. The videos at each of the addresses are playable by us and several colleagues who also checked.*

I think the visualization you propose is very clear.
However, I have the impression that it is difficult to understand, so I expect detailed explanations.
For example, detailed explanations of Figure 3 e and Figure 4 e would be helpful.
I also think that explanations of the data in Figure 3 and Figure 4 would make it even easier to understand.

*TODO: expand explanations*

For small details, I think it is necessary to check page 5, line 38 and page 11, line 29.

*TODO: not sure what the problem is here, maybe some typo*

### 2023-11-06 review/cheem_ Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections/sn-article.tex

\documentclass[12pt]{sn-jnl}

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}

\usepackage[utf8]{inputenc}

\author{
Nicholas Spyrison~\orcidlink{0000-0002-8417-0212}\\Monash
University \And Dianne Cook~\orcidlink{0000-0002-3813-7155}\\Monash
University \And Przemyslaw
Biecek~\orcidlink{0000-0001-8423-1823}\\University of Warsaw\\
Warsaw University of Technology
}
\title{}

\Plainauthor{Nicholas Spyrison, Dianne Cook, Przemyslaw Biecek}

\Abstract{
The increased predictive power of machine learning models comes at the
cost of increased complexity and loss of interpretability, particularly
in comparison to parametric statistical models. This trade-off has led
to the emergence of eXplainable AI (XAI) which provides methods, such as
local explanations (LEs) and local variable attributions (LVAs), to shed
light on how a model use predictors to arrive at a prediction. These
provide a point estimate of the linear variable importance in the
vicinity of a single observation. However, LVAs tend not to effectively
handle association between predictors. To understand how the interaction
between predictors affects the variable importance estimate, we can
convert LVAs into linear projections and use the radial tour. This is
also useful for learning how a model has made a mistake, or the effect
of outliers, or the clustering of observations. The approach is
illustrated with examples from categorical (penguin species, chocolate
types) and quantitative (soccer/football salaries, house prices)
response models. The methods are implemented in the R package cheem,
available on CRAN.
}

\Keywords{explainable artificial intelligence, nonlinear model
interpretability, visual analytics, local explanations, grand
tour, radial tour}
\Plainkeywords{explainable artificial intelligence, nonlinear model
interpretability, visual analytics, local explanations, grand
tour, radial tour}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    Nicholas Spyrison\\
    Monash University\\
    Dept of Human Centred Computing\\
Faculty of Information Technology\\
  E-mail: \email{spyrison@gmail.com}\\
  URL: \url{https://nspyrison.netlify.app}\\~\\
      Dianne Cook\\
    Monash University\\
    Dept of Econometrics and Business Statistics\\

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{amsmath} \usepackage{datetime}

@@

@@

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{./figures/shap_distr_bd}

\caption[Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors]{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors. A star offensive and defensive player are compared, L. Messi and V. van Dijk, respectively. Panel (a) shows breakdown plots of three sequences of the variables. The sequence of the variables impacts the magnitude of their attribution. Panel (b) shows the distribution of attribution for each variable across 25 sequences of predictors, with the mean displayed as a dot for each player. Reaction skills are important for both players. Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
\end{figure}
\end{CodeChunk}

@@

There are numerous R packages currently available that provide functions
for computing SHAP values, including \texttt{fastshap} \citep{fastshap},
\texttt{kernelshap} \citep{kernelshap}, \texttt{shapr} \citep{shapr},
\texttt{shapviz} \citep{shapviz}, \texttt{PPtreeregViz}
\citep{PPtreeregViz}, \texttt{ExplainPrediction}
\citep{ExplainPrediction}, \texttt{flashlight} \citep{flashlight}, and
the package \texttt{DALEX} has many resources \citep{biecek_dalex_2018}.
There are many more packages only available through Github, like
\texttt{treeshap} \citep{kominsarczyk_treeshap_2021} that is used for
this work. \citet{molnar2022} provides good explanations of the
different methods and how to apply them to different models.

@@

A version of the manual tour called a \emph{radial tour} is implemented
in \citet{spyrison_spinifex_2020} and forms the basis of this new work.
In a radial tour, the selected variable can change its magnitude of
contribution but not its angle; it must move along the direction of its
original contribution. The implementation allows for pre-computation and
interactive re-calculation to focus on a different variable.

\begin{CodeChunk}
\begin{figure}

@@

The visualization is composed of the first two principal components of
the data (left) and the attribution (middle) spaces. These single 2D
projections will not reveal all of the structure of higher-dimensional
space, but they are helpful visual summaries. In addition, a plot of the
observed against predicted response values is also provided (Figures
\ref{fig:classificationcase}b, \ref{fig:regressioncase}a) to help
identify observations poorly predicted by the model. For classification
tasks, color indicates the predicted class and misclassified
observations are circled in red. Linked brushing between the plots is
provided, and a tabular display of selected points helps to facilitate
the exploration of the spaces and the model (shown in Figures
\ref{fig:regressioncase}d).

@@

@@

A bar chart is overlaid to represent the projection shown in the radial
tour on the right. It starts from the SHAP values of the PI, but if the
user changes the projection the length of these bars will reflect this
change. (The PI is interactively selected by clicking on a point in the
global view). By scaling the SHAP value it becomes an (attribution)
projection.

The attribution projection of the PI is the initial 1D basis in a radial
tour, displayed as a density plot for a categorical response (Figure
\ref{fig:classificationcase}f) and as scatterplots for a quantitative
response (Figure \ref{fig:regressioncase}f). The PI and CI are indicated
by vertical dashed and dotted lines, respectively. The radial tour
varies the contribution of the selected variable. This is viewed as an
animation of the projections from many intermediate bases. Doing so
tests the sensitivity of structure (class separation or strength of
relationship) to the variable's contribution. For classification, if the
separation between classes diminishes when the variable contribution is
reduced, this suggests that the variable is important for class
separation. For regression, if the relationship scatterplot weakens when
the variable contribution is reduced, indicating that the variable is
important for accurately predicting the response.

@@

\begin{CodeChunk}
\begin{figure}

@@

\caption[Overview of the cheem viewer for classification tasks (categorical response)]{Overview of the cheem viewer for classification tasks (categorical response). Global view inputs, (a), set the PI, CI, and color statistic. Global view, (b) PC1 by PC2 approximations of the data- and attribution-space. (c) prediction by observed $y$ (visual of the confusion matrix for classification tasks). Points are colored by predicted class, and red circles indicate misclassified observations. Radial tour inputs (d) select variables to include and which variable is changed in the tour. (e) shows a parallel coordinate display of the distribution of the variable attributions while bars depict contribution for the current basis. The black bar is the variable being changed in the radial tour. Panel (f) is the resulting data projection indicated as density in the classification case.}\label{fig:classificationcase}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/app_regression_interactions}

\caption[Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables]{Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables. Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values. Four selected points are highlighted in the PC spaces and tabularly displayed. Coloring on a statistic (c) highlights the structure organized in the attribution space. Interactive tabular display (d) populates when observations are selected. Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the variable attribution from all observations, and horizontal bars show the contribution to the current basis. Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed $y$ and residuals (middle and right).}\label{fig:regressioncase}
\end{figure}
\end{CodeChunk}

@@

A tooltip displays the observation number/name and classification
information while the cursor hovers over a point. Linked brushing allows
the selection of points (left click and drag) where those points will be
highlighted across plots (Figure \ref{fig:classificationcase}a \& b).
The information corresponding to the selected points is populated on a
dynamic table (Figure \ref{fig:classificationcase}d). These interactions
aid the exploration of the spaces and, finally, the identification of
primary and comparison observations.

@@

\begin{itemize}
\tightlist
\item
  \textbf{Data:} predictors and response are unscaled complete numerical
  matrix. Most models and local explanations are scale-invariant. Keep
  the normality assumptions of the model in mind.
\item
  \textbf{Model:} any model and compatible explanation could be explored
  with this method. Currently, random forest models are applied via the
  package \textbf{randomForest} \citep{liaw_classification_2002},
  compatibility tree SHAP. Modest hyperparameters are used, namely: 125
  trees, the number of variables at each split, mtry = \(\sqrt{p}\) or
  \(p/3\) for classification and regression, and minimum size of
  terminal nodes \(max(1, n/500)\) or \(max(5, n/500)\) for
  classification and regression.
\item
  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each}
  observation using the package \textbf{treeshap}
  \citep{kominsarczyk_treeshap_2021}. We opt to find the attribution of
  each observation in the training data and not fit to fit variable
  interactions.
\item
  \textbf{Cheem viewer:} after the model and full explanation space are
  calculated, each variable is scaled by standard deviations away from
  the mean to achieve common support for visuals. Statistics for mapping
  to color are computed on the scaled spaces.
\end{itemize}

@@

\caption[Examining the SHAP values for a random forest model classifying Palmer penguin species]{Examining the SHAP values for a random forest model classifying Palmer penguin species. The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b). The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right). (The animation of the radial tour is at https://vimeo.com/666431172.)}\label{fig:casepenguins}
\end{figure}
\end{CodeChunk}

@@

\caption[Checking what is learned from the cheem viewer]{Checking what is learned from the cheem viewer. This is a plot of flipper length (`fl`) and bill length (`bl`), where an asterisk highlights the PI. A Gentoo (purple) misclassified as a Chinstrap (orange). The PI has an unusually small `fl` length which is why it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
\end{figure}
\end{CodeChunk}

@@

\caption[Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green)]{Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green). From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the LVA does not help understand the prediction for this observation. So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber. When Sugar is zero, Fiber contributes strongly toward the left. In this view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber. This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate. (A video of the tour animation can be found at https://vimeo.com/666431143.)}\label{fig:casechocolates}
\end{figure}
\end{CodeChunk}

@@

\caption[Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange)]{Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange). In the attribution projection, the PI could be either milk or dark. Sodium and Fiber have the largest differences in attributed variable importance, with low values relative to other milk chocolates. The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied. If Sodium had a larger contribution to the prediction (like in this view). the PI would look more like other milk chocolates. (A video of the tour animation can be found at https://vimeo.com/666431148.)}\label{fig:casechocolatesinverse}
\end{figure}
\end{CodeChunk}

@@

\caption[Exploring the wages relative to skill measurements in the FIFA 2020 data]{Exploring the wages relative to skill measurements in the FIFA 2020 data. Star offensive player (L. Messi) is the PI, and he is compared with a top defensive player (V. van Dijk). The attribution projection is shown on the left, and it can be seen that this combination of variables produces a view where Messi has very high predicted (and observed) wages. Defense (`def`) is the chosen variable to vary. It starts very low, and Messi's predicted wages decrease dramatically as its contribution increases (right plot). The increased contribution in defense comes at the expense of offensive and reaction skills. The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the LVA. (A video of the animated radial tour can be found at https://vimeo.com/666431163.)}\label{fig:casefifa}
\end{figure}
\end{CodeChunk}

@@

\caption[Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data]{Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data. The LVA indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable. The PI has a higher predicted value than the CI in the attribution projection. Reducing the contribution of Lot Area brings these two prices in line. This suggests that if the model did not value Lot Area so highly for this observation, then the observed sales price would be quite similar. That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price. (A video showing the animation is at https://vimeo.com/666431134.)}\label{fig:caseames}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> install.packages("cheem", dependencies = TRUE)
R> library("cheem")
R> run_app()
\end{CodeInput}
\end{CodeChunk}

Alternatively,

\begin{itemize}
\tightlist
\item
  A version of the cheem viewer shiny app can be directly accessed at
  \url{https://ebsmonash.shinyapps.io/cheem/}.
\item
  The development version of the package is available at
  \url{https://github.com/nspyrison/cheem}, and
\item
  Documentation of the package can be found at
  \url{https://nspyrison.github.io/cheem/}.
\end{itemize}

@@

\hypertarget{acknowledgments}{%
\subsection*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{subsection}{Acknowledgments}

@@

\renewcommand\refname{References}
\bibliography{paper.bib}

>>>>

%Version 2.1 April 2023
% See section 11 of the User Manual for version history
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing

%%=======================================================%%
%% to print line numbers in the margin use lineno option %%
%%=======================================================%%

%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%======================================================%%
%% to compile with pdflatex/xelatex use pdflatex option %%
%%======================================================%%

%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%Note: the following reference styles support Namedate and Numbered referencing. By default the style follows the most common style. To switch between the options you can add or remove Numbered in the optional parenthesis. 
%%The option is available for: sn-basic.bst, sn-vancouver.bst, sn-chicago.bst, sn-mathphys.bst. %

%%\documentclass[sn-nature]{sn-jnl}% Style for submissions to Nature Portfolio journals
%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
%%\documentclass[sn-mathphys,Numbered]{sn-jnl}% Math and Physical Sciences Reference Style
%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
%%\documentclass[sn-vancouver,Numbered]{sn-jnl}% Vancouver Reference Style
%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style 
\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
%%\documentclass[default]{sn-jnl}% Default
%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout

%%%% Standard Packages
%%<additional latex packages if required can be included here>

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published 
%%%%  by Springer Nature. The guidance has been prepared in partnership with 
%%%%  production teams to conform to Springer Nature technical requirements. 
%%%%  Editorial and presentation requirements differ among journal portfolios and 
%%%%  research disciplines. You may find sections in this template are irrelevant 
%%%%  to your work and are empowered to omit any such section if allowed by the 
%%%%  journal you intend to submit to. The submission guidelines and policies 
%%%%  of the journal take precedence. A detailed User Manual is available in the 
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

%\jyear{2021}%

%% as per the requirement new theorem styles can be included as shown below
\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}% 
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads

@@

\title[Article Title]{Article Title}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate} 
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\title{Exploring Local Explanations of Nonlinear Models Using Animated Linear
Projections}

\author*[1]{\fnm{Nicholas} \sur{Spyrison}}\email{spyrison@gmail.com}

\author[2]{\fnm{Dianne} \sur{Cook}}\email{dicook@monash.edu}

\author[3]{\fnm{Przemyslaw} \sur{Biecek}}\email{przemyslaw.biecek@pw.edu.pl}

\affil*[1]{\orgdiv{Faculty of Information Technology}, \orgname{Monash
University}, \orgaddress{\street{Wellington Rd}, \city{Clayton}, \postcode{VIC 3800}, \country{Australia}}}

\affil[2]{\orgdiv{Dept of Econometrics and Business Statistics}, \orgname{Monash University}, \orgaddress{\street{Wellington Rd}, \city{Clayton}, \postcode{VIC 3800}, \country{Australia}}}

\affil[3]{\orgdiv{Faculty of Mathematics and Information Science}, \orgname{Warsaw University of Technology}, \orgaddress{\street{Koszykowa 75}, \city{Warsaw}, \postcode{00-662}, \country{Poland}}}

%%==================================%%
%% sample for unstructured abstract %%
%%==================================%%

\abstract{The increased predictive power of machine learning models comes at the cost of increased complexity and loss of interpretability, particularly in comparison to parametric statistical models. This trade-off has led to the emergence of eXplainable AI (XAI) which provides methods, such as local explanations (LEs) and local variable attributions (LVAs), to shed light on how a model use predictors to arrive at a prediction. These provide a point estimate of the linear variable importance in the vicinity of a single observation. However, LVAs tend not to effectively handle association between predictors. To understand how the interaction between predictors affects the variable importance estimate, we can convert LVAs into linear projections and use the radial tour. This is also useful for learning how a model has made a mistake, or the effect of outliers, or the clustering of observations. The approach is illustrated with examples from categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) response models. The methods are implemented in the R package cheem, available on CRAN.}

%%================================%%
%% Sample for structured abstract %%
%%================================%%

\keywords{explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour}

%%\pacs[JEL Classification]{D8, H51}

%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}

\maketitle

@@

\begin{figure}[h!]
{\centering \includegraphics[width=0.85\linewidth]{figures/shap_distr_bd} 
}

\caption{Illustration of SHAP values for a random forest model FIFA 2020 player wages from nine skill predictors. A star offensive and defensive player are compared, L. Messi and V. van Dijk, respectively. Panel (a) shows breakdown plots of three sequences of the variables. The sequence of the variables impacts the magnitude of their attribution. Panel (b) shows the distribution of attribution for each variable across 25 sequences of predictors, with the mean displayed as a dot for each player. Reaction skills are important for both players. Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi.}\label{fig:shapdistrbd}
\end{figure}

@@

There are numerous R packages currently available on CRAN that provide functions
for computing SHAP and other LVA values, including \texttt{treeshap} \citep{kominsarczyk_treeshap_2023}, \texttt{fastshap} \citep{fastshap},
\texttt{kernelshap} \citep{kernelshap}, \texttt{shapr} \citep{shapr},
\texttt{shapviz} \citep{shapviz}, \texttt{PPtreeregViz}
\citep{PPtreeregViz}, \texttt{ExplainPrediction}
\citep{ExplainPrediction}, \texttt{flashlight} \citep{flashlight}, and
the package \texttt{DALEX} has many resources \citep{biecek_dalex_2018}.
\citet{molnar2022} provides good explanations of the
different methods and how to apply them to different models.

@@

A version of the manual tour called a \emph{radial tour} is implemented
in \citet{spyrison_spinifex_2020} and forms the basis of this new work.
In a radial tour, the selected variable can change its magnitude of
contribution but not its angle; it must move along the direction of its
original contribution. The implementation allows for pre-computation and
interactive re-calculation to focus on a different variable. In this work, the radial tour allows us to explore the sensitivity of LVA to the prediction of a model.

\begin{figure}[h!]

@@

The visualization is composed of the first two principal components of
the data (left) and the attribution (middle) spaces. These single 2D
projections will not reveal all of the structure of higher-dimensional
space, but they are helpful visual summaries. In addition, a plot of the
observed against predicted response values is also provided (Figures
\ref{fig:classificationcase}c, \ref{fig:regressioncase}b) to help
identify observations poorly predicted by the model. For classification
tasks, color indicates the predicted class and misclassified
observations are circled in red. Linked brushing between the plots is
provided (click and drag), and a tabular display of selected points helps to facilitate
the exploration of the spaces and the model (shown in Figure
\ref{fig:regressioncase}d).

@@

HERE

@@

A bar chart is overlaid to represent the projection shown in the radial
tour on the right. It starts from the SHAP values of the PI, but if the
user changes the projection the length of these bars will reflect this
change. By scaling the SHAP value it becomes an (attribution)
projection.

The attribution projection of the PI is the initial 1D basis in a radial
tour, displayed as a density plot for a categorical response (Figure
\ref{fig:classificationcase}f) and as scatterplots for a quantitative
response (Figure \ref{fig:regressioncase}f). The PI and CI are indicated
by vertical dashed and dotted lines, respectively. The radial tour
varies the contribution of the selected variable. This is viewed as an
animation of the projections from many intermediate bases. Doing so
tests the sensitivity of structure (class separation or strength of
relationship) to the variable's contribution. The CI attribution of 
the CI does not impact the bases but it highlighted from context. 
For classification, if the separation between classes diminishes when the variable contribution is
reduced, this suggests that the variable is important for class
separation. For regression, if the relationship scatterplot weakens when
the variable contribution is reduced, indicating that the variable is
important for accurately predicting the response.

@@

\begin{figure}[h!]

@@

\caption{Overview of the cheem viewer for classification tasks (categorical response). Global view inputs, (a), set the PI, CI, and color statistic. Global view, (b) PC1 by PC2 approximations of the data- and attribution-space. (c) prediction by observed $y$ (visual of the confusion matrix for classification tasks). Points are colored by predicted class, and red circles indicate misclassified observations. Radial tour inputs (d) select variables to include and which variable is changed in the tour. (e) shows a parallel coordinate display of the distribution of the variable attributions while bars depict contribution for the current basis. The black bar is the variable being changed in the radial tour. Panel (f) is the resulting data projection indicated as density in the classification case.}\label{fig:classificationcase}
\end{figure}

@@

\begin{figure}[h!]

{\centering \includegraphics[width=1\linewidth]{./figures/app_regression_interactions} 
}

\caption{Overview of the cheem viewer for regression tasks (quantitative response) and illustration of interactive variables. Panel (a) PCA of the data- and attributions- spaces and the (b) residual plot, predictions by observed values. Four selected points are highlighted in the PC spaces and tabularly displayed. Coloring on a statistic (c) highlights the structure organized in the attribution space. Interactive tabular display (d) populates when observations are selected. Contribution of the 1D basis affecting the horizontal position (e) parallel coordinate display of the variable attribution from all observations, and horizontal bars show the contribution to the current basis. Regression projection (f) uses the same horizontal projection and fixes the vertical positions to the observed $y$ and residuals (middle and right).}\label{fig:regressioncase}
\end{figure}

@@

A tooltip displays the observation number/name and classification
information while the cursor hovers over a point. Linked brushing allows
the selection of points (left click and drag) where those points will be
highlighted across plots (Figure \ref{fig:regressioncase}a \& b).
The information corresponding to the selected points is populated on a
dynamic table (Figure \ref{fig:regressioncase}d). These interactions
aid the exploration of the spaces and, finally, the identification of
primary and comparison observations.

@@

\begin{itemize}
\tightlist
\item
  \textbf{Data:} predictors and response are unscaled complete numerical
  matrix. Most models and local explanations are scale-invariant. Keep
  the normality assumptions of the model in mind.
\item
  \textbf{Model:} any model and compatible explanation could be explored
  with this method. Currently, random forest models are applied via the
  package \textbf{randomForest} \citep{liaw_classification_2002},
  compatibility tree SHAP. Modest hyperparameters are used. Namely, 
  classification models use 125 trees, number of variables at each 
  split (mtry) of \(\sqrt{p}\), and minimum terminal node size of 
  \(max(1, n/500)\). While regression models use 125 tree, \(p/3\) variables at split, and 
  \(max(5, n/500)\) minimum terminal node size.
\item
  \textbf{Local explanation:} Tree SHAP is calculated for \emph{each}
  observation using the package \textbf{treeshap}
  \citep{kominsarczyk_treeshap_2023}. We opt to find the attribution of
  each observation in the training data and not fit to fit variable
  interactions.
\item
  \textbf{Cheem viewer:} after the model and full explanation space are
  calculated, each variable is scaled by standard deviations away from
  the mean to achieve common support for visuals. Statistics for mapping
  to color are computed on the scaled spaces.
\end{itemize}

@@

\caption{Examining the SHAP values for a random forest model classifying Palmer penguin species. The PI is a Gentoo (purple) penguin that is misclassified as a Chinstrap (orange), marked as an asterisk in (a) and the dashed vertical line in (b). The radial view shows varying the contribution of `fl` from the initial attribution projection (b, left), which produces a linear combination where the PI is more probably (higher density value) a Chinstrap than a Gentoo (b, right). (The animation of the radial tour is at https://vimeo.com/666431172.)}\label{fig:casepenguins}
\end{figure}

@@

\caption{Checking what is learned from the cheem viewer. This is a plot of flipper length (`fl`) and bill length (`bl`), where an asterisk highlights the PI. A Gentoo (purple) misclassified as a Chinstrap (orange). The PI has an unusually small `fl` length which is why it is confused with a Chinstrap.}\label{fig:casepenguinsblfl}
\end{figure}

@@

\caption{Examining the LVA for a PI which is dark (orange) chocolate incorrectly predicted to be milk (green). From the attribution projection, this chocolate correctly looks more like dark than milk, which suggests that the LVA does not help understand the prediction for this observation. So, the contribution of Sugar is varied---reducing it corresponds primarily with an increased magnitude from Fiber. When Sugar is zero, Fiber contributes strongly toward the left. In this view, the PI is closer to the bulk of the milk chocolates, suggesting that the prediction put a lot of importance on Fiber. This chocolate is a rare dark chocolate without any Fiber leading to it being mistaken for a milk chocolate. (A video of the tour animation can be found at https://vimeo.com/666431143.)}\label{fig:casechocolates}
\end{figure}

@@

\caption{Examining the LVA for a PI which is milk (green) chocolate incorrectly predicted to be dark (orange). In the attribution projection, the PI could be either milk or dark. Sodium and Fiber have the largest differences in attributed variable importance, with low values relative to other milk chocolates. The lack of importance attributed to these variables is suspected of contributing to the mistake, so the contribution of Sodium is varied. If Sodium had a larger contribution to the prediction (like in this view). the PI would look more like other milk chocolates. (A video of the tour animation can be found at https://vimeo.com/666431148.)}\label{fig:casechocolatesinverse}
\end{figure}

@@

\caption{Exploring the wages relative to skill measurements in the FIFA 2020 data. Star offensive player (L. Messi) is the PI, and he is compared with a top defensive player (V. van Dijk). The attribution projection is shown on the left, and it can be seen that this combination of variables produces a view where Messi has very high predicted (and observed) wages. Defense (`def`) is the chosen variable to vary. It starts very low, and Messi's predicted wages decrease dramatically as its contribution increases (right plot). The increased contribution in defense comes at the expense of offensive and reaction skills. The interpretation is that Messi's high wages are most attributable to his offensive and reaction skills, as initially provided by the LVA. (A video of the animated radial tour can be found at https://vimeo.com/666431163.)}\label{fig:casefifa}
\end{figure}

@@

\caption{Exploring an observation with an extreme residual as the PI in relation to an observation with an accurate prediction for a similarly priced house in a random forest fit to the Ames housing data. The LVA indicates a sizable attribution to Lot Area (`LtA`), while the CI has minimal attribution to this variable. The PI has a higher predicted value than the CI in the attribution projection. Reducing the contribution of Lot Area brings these two prices in line. This suggests that if the model did not value Lot Area so highly for this observation, then the observed sales price would be quite similar. That is, the large residual is due to a lack of factoring in the Lot Area for the prediction of PI's sales price. (A video showing the animation is at https://vimeo.com/666431134.)}\label{fig:caseames}
\end{figure}

@@

\begin{lstlisting}
R> install.packages("cheem", dependencies = TRUE)
R> library("cheem")
R> run_app()
\end{lstlisting}

Alternatively, a version of the cheem viewer shiny app can be directly accessed at \url{https://ebsmonash.shinyapps.io/cheem/}, the development version of the package is available at \url{https://github.com/nspyrison/cheem}, and documentation of the package can be found at \url{https://nspyrison.github.io/cheem/}.

@@

\backmatter

\bmhead{Acknowledgments}

@@

%%=============================================%%
%% For submissions to Nature Portfolio Journals %%
%% please use the heading ``Extended Data''.   %%
%%=============================================%%

%%=============================================================%%
%% Sample for another appendix section			       %%
%%=============================================================%%

%% \section{Example of another appendix section}\label{secA2}%
%% Appendices may be used for helpful, supporting or essential material that would otherwise 
%% clutter, break up or be distracting to the text. Appendices can consist of sections, figures, 
%% tables and equations etc.
